# Analysis of the Major Storm and Weather events in the US from 1950 to 2011.

In this project we aim to analyse the U.S. National Oceanic and Atmospheric Administration (NOAA) database from 1950 to 2011 and adress the following questions: i) Which events are the most harmful to population health? ii) Which events have the gratest economic consequences?. Our results showed that tornados accounted the highest Fatalities and Injuries in the last 61 years in the US ~ 37 % of the Fatalities and 65% of the injuries. 
Hurricanes and Typhoons, followed by Floods were the most expensive events to the US, accounting for 37, 36.4 and 10.5 % of the total expenses. Crop related expenses were more than 4-fold the Properties expenses.

## Data Processing
``` {r Data Processing}
setwd("C:/Users/Marcelo/Desktop/Data Science/Data-Science-Specialization/Reproducible Research/")
destfile="./NOAA.csv.bz2" 
fileURL <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"   
download.file(fileURL ,destfile,method="auto")

NOAA <- read.csv("NOAA.csv.bz2", header = TRUE, stringsAsFactors = FALSE)

head(NOAA)
```

There are several variables that will not be used in this project, so I will subset the loaded data set.

 The variables we will use are:  

 EVTYPE as a measure of event type (e.g. tornado, flood, etc.)  
 FATALITIES as a measure of harm to human health  
 INJURIES as a measure of harm to human health  
 PROPDMG as a measure of property damage and hence economic damage in USD  
 PROPDMGEXP as a measure of magnitude of property damage (e.g. thousands, millions USD, etc.)
 CROPDMG as a measure of crop damage and hence economic damage in USD  
 CROPDMGEXP as a measure of magnitude of crop damage (e.g. thousands, millions USD, etc.)
 
``` {r Filtering Data}
vars <- c("EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP")

NOAA_dat <- NOAA[vars]

# Check for missing values NA's
sum(is.na(NOAA_dat)) # no NA's
```

Economic values are stored  in separated columns described with letters:  
H = Hundreds, K = Thousands, M = Millions, B = Billions. Calculations of the   total value could be trick:

``` {r chunk 3}
exp <- function(e){
  if(e %in% c("h","H"))
    return(2)
  else if(e %in% c("k","K"))
    return(3)
  else if(e %in% c("m","M"))
    return(6)
  else if(e %in% c("b","B"))
    return(9)
  else if(!is.na(as.numeric(e)))
    return(as.numeric(e))
  else if (e %in% c("", "-", "?", "+"))
    return(0)
  else{
    stop("invalid Value.")
  }
}

```

Assinging values to Property damages and crop damages

``` {r chunk 4}
propExp <- sapply(NOAA_dat$PROPDMGEXP, FUN = exp)
NOAA_dat$PropDamage <- NOAA_dat$PROPDMG*(10**propExp)
cropExp <-  sapply(NOAA_dat$CROPDMGEXP, FUN = exp)
NOAA_dat$CropDamage <- NOAA_dat$CROPDMG*(10**propExp)
```

Removing unused columns

``` {r chunk 5}
NOAA_dat$PROPDMG <- NULL
NOAA_dat$PROPDMGEXP <- NULL
NOAA_dat$CROPDMG <- NULL
NOAA_dat$CROPDMGEXP <- NULL
```

## Results

### Fatalities and Injuries

```{r chunk6}
library(dplyr)
library(ggplot2)
library(gridExtra)

fatality_events <- NOAA_dat %>%
  group_by(EVTYPE) %>%
  summarise(Fatalities = sum(FATALITIES),
            Percentage = round(Fatalities/sum(NOAA_dat$FATALITIES)*100,2))%>%
  arrange(desc(Fatalities)) %>%
  head(10)

injury_events <- NOAA_dat %>%
  group_by(EVTYPE) %>%
  summarise(Injuries = sum(INJURIES),
            Percentage = round(Injuries/sum(NOAA_dat$INJURIES)*100,2))%>%
  arrange(desc(Injuries)) %>%
  head(10)

p1 <- ggplot(data = fatality_events, aes(x = reorder(EVTYPE, Fatalities), y = Fatalities,
                           fill = Fatalities, label = round(Percentage,2)))+
  geom_bar(stat = "identity")+
  coord_flip()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Event Type")+
  ylab("Number of Fatalities")+
  ggtitle("Top 10 Events with Higest Fatalities in the US from 1950 to 2011")+
  theme(legend.position="none")+
  geom_label(aes(fill = Fatalities), colour = "white", fontface = "bold")


p2 <- ggplot(data = injury_events, aes(x = reorder(EVTYPE, Injuries), y = Injuries,
                                         fill = Injuries, label = round(Percentage,2)))+
  geom_bar(stat = "identity")+
  coord_flip()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Event Type")+
  ylab("Number of Injuries")+
  ggtitle("Top 10 Events with Higest Injuries in the US from 1950 to 2011")+
  theme(legend.position="none")+
  geom_label(aes(fill = Injuries), colour = "white")

grid.arrange(p1, p2, nrow = 2)
```

Tornados were responsible for the highest Fatalities and Injuries in the last 61 years in the US reponsible for ~ 37 % of the Fatalities and 65% of the injuries.

In the fatalitites case, Excessive heat, flash food, heat and lightning complete the top 5 events accounting for more than 67 % of the fatalities during the period.

### Economic Impact

```{r chunk 7}

NOAA_dat$TotalDMG <- NOAA_dat$PropDamage + NOAA_dat$CropDamage

TotalDMG <- NOAA_dat %>%
  group_by(EVTYPE) %>%
  summarise(Total = sum(TotalDMG),
            Percentage = round(Total/sum(NOAA_dat$TotalDMG)*100,2))%>%
  arrange(desc(Total)) %>%
  head(10)

ggplot(data = TotalDMG, aes(x = reorder(EVTYPE, -Total), y = Total/1000000000,
                                       fill = Total, label = round(Percentage,2)))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Event Type")+
  ylab("Expense in U$ / billions")+
  ggtitle("Top 10 Events Most Expensive events in the US from 1950 to 2011")+
  theme(legend.position="none")+
  geom_label(aes(fill = Total), colour = "white")

proDMG <- NOAA_dat %>%
  group_by(EVTYPE) %>%
  summarise(Total = sum(PropDamage),
            Percentage = round(Total/sum(NOAA_dat$PropDamage)*100,2))%>%
  arrange(desc(Total)) %>%
  head(10)

ggplot(data = proDMG, aes(x = reorder(EVTYPE, -Total), y = Total/1000000000,
                                  fill = Total, label = round(Percentage,2)))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Event Type")+
  ylab("Expense in U$ / billions")+
  ggtitle("Top 10 Most Expensive Properties Events in the US from 1950 to 2011")+
  theme(legend.position="none")+
  geom_label(aes(fill = Total), colour = "white")

cropDMG <- NOAA_dat %>%
  group_by(EVTYPE) %>%
  summarise(Total = sum(CropDamage),
            Percentage = round(Total/sum(NOAA_dat$CropDamage)*100,2))%>%
  arrange(desc(Total)) %>%
  head(10)

ggplot(data = cropDMG, aes(x = reorder(EVTYPE, -Total), y = Total/1000000000,
                                fill = Total, label = round(Percentage,2)))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  xlab("Event Type")+
  ylab("Expense in U$ / billions")+
  ggtitle("Top 10 Most Expensive Crop Events in the US from 1950 to 2011")+
  theme(legend.position="none")+
  geom_label(aes(fill = Total), colour = "white")

```

Considering the economic impact during the 61 years period, Hurricanes and Typhoons, followed by Floods
were the most expensive events to the US, accounting for 37, 36.4 and 10.5 % of the total expenses so far, respectively.

Spliting our analysis into Properties and Crop impacts, it can be observed that the events that affected Properties were Flood, Hurricane and Tornados, representing more than 50% of the expenses. On the other hand, the events that affected the Crop were mainly Hurricane and Thypoon, representing more than 96% of the expenses.

It worth mentioning that the economic damages related to the Crop were more than 4 times the economic loss related to the Properties. 

# Conclusion

Tornados accounted the highest Fatalities and Injuries in the last 61 years in the US ~ 37 % of the Fatalities and 65% of the injuries. 
Hurricanes and Typhoons, followed by Floods were the most expensive events to the US, accounting for 37, 36.4 and 10.5 % of the total expenses. Crop related expenses were more than 4-fold the Properties expenses.
