---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Statistical Inference Project
Marcelo Nobrega

# Part 1: Simulation Exercise Instructions

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

## 1. Show the sample mean and compare it to the theoretical mean of the distribution.

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

```{r data}
set.seed(2017)
lambda <- 0.2
n <- 40 # number of exponentials
nosim <- 1000 # number of simulations

# The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter
exp_sim <- matrix(rexp(n * nosim, rate = lambda), nosim) # simulating data
means <- rowMeans(exp_sim) #calculating the means
mean_sim <- mean(means)
mean_sim
```
Theoretical mean
```{r}
T_mean <- 1/ lambda
T_mean
```

```{r}
hist(means, col = "white", main = "rexp mean distribution x Theoretical Mean", breaks = 40)
abline(v=mean_sim, lwd = "4", col = "blue")
abline(v=5.0, lwd = "4", col = "orange")
text(7, 40, paste("Simulated mean = ", round(mean_sim, 2), "\n Theoretical mean = 5.0"), col = "black")
```


The simulated mean is 4.9828 and the theoretical mean is 5, showing that the center of distributuion of the simulated data is very close to the theoretical center of distribution. 

## 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r}
Expvar <- var(means)
Expvar

# Theoretical Var
T_var <- (1/lambda)^2/n
T_var
```
Variance of the sample means is 0.6035 while the theoretical variance is 0.625, very close values. 

## 3. Show that the distribution is approximately normal.

```{r echo=FALSE}
library(ggplot2)
T_sd <- 1/(lambda * sqrt(n))
data <- data.frame(means)
ggplot(data, aes (x = means))+
  geom_histogram(aes(y = ..density..), colour = "black", fill = "lightgreen")+
  ggtitle("Distribution of 40 Simulated averages x Theoretical Normal distribution")+
  xlab("Mean")+
  ylab("Density")+
  geom_vline(aes(xintercept = mean_sim), colour = "blue")+
  geom_vline(aes(xintercept = T_mean), colour = "red")+
  stat_function(fun = dnorm, args = list(mean = mean_sim, sd = sd(means)), colour = "blue", size = 1.0)+
  stat_function(fun = dnorm, args = list(mean = T_mean, sd = T_sd), colour = "red", size = 1.0)
```

As presented the the simulatd mean and the theoretical one are vey similar. Additionally, the simulated data has a distribution that resembles a normal distrib (blue trace), presenting a great overlap with the normal distribution (red trace).
