---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Statistical Inference Project
Marcelo Nobrega

The project consists of two parts:

2. Basic inferential data analysis.

# Part 2: Basic Inferential Data Analysis Instructions

## 1. Provide a basic summary of the data.

We're going to analyze the ToothGrowth data in the R datasets package.

Loading libraries, the ToothGrowth data and performing some basic exploratory data analyses.
```{r}
library(ggplot2)
data("ToothGrowth")
# Some basic exploratory data analyses
summary(ToothGrowth)
head(ToothGrowth)
tail(ToothGrowth)
str(ToothGrowth)
```
The ToothGrowth data consists in the length of odontoblasts (cells responsible for tooth growth)
in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) 
by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

60 observations and 3 variables. 
len: Tooth length
Supp: Supplement type (VC or OJ)
dose: Dose in miligrams/day

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
ggplot(ToothGrowth, aes(x = dose, y = len))+
  geom_boxplot(aes(fill = dose))+
  facet_grid(~ supp)+
  xlab("Dose (mg / day)")+
  ylab("Tooth length")+
  ggtitle("Tooth Length per Dose by Delivery Method")

```

The larger is the dosage for both delivery methods longre is th tooth.

## 2. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r}
unique(ToothGrowth$dose)
```

There are 3 levels which can be understood as 3 doseage groups, 0.1, 1.0 and 2.0 mg / day. I will compare the tooth growth by each doseage group and delivery method.

Comparing tooth groth by Delivery method using t-test.
```{r}
t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose == 0.5,])
```

```{r}
t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose == 1,])
```

```{r}
t.test(len ~ supp, data = ToothGrowth[ToothGrowth$dose == 2,])
```

## 3. State your conclusions and the assumptions needed for your conclusions.
It was assumed that the sample data was paired.

Conclusions: 
Dosage equal to 0.5 mg/day: OJ had higher mean tooth length than VC, 13.23 vs. 7.98. P-value is 0.006 (< 0.05), so for the same dosage the different delivery methods have effect on the tooth length.

Dosage equal to 1.0 mg/day: OJ had higher mean tooth length than VC, 22.70 vs. 16.77. P-value is 0.001 (< 0.05). The different delivery methods have effect on the tooth length.

Dosage equal to 2.0 mg/day: Both delivery methods, OJ and VC, had similar average tooth length, 26.06 vs. 26.14. P-value is 0.963 (> 0.05), and the confidence interval conains zero, so the null hypothesis can't be rejected and the different delivery methods have no effect on the tooth length.



