T test in R

```{r libraries}
library(UsingR); data(father.son)
```

```{r test}
# comparing son's heights and fathers heights
t.test(father.son$sheight-father.son$fheight)

```

ChickWeight data - two group intervals

```{r chick}
library(datasets); data(ChickWeight); library(reshape2)

# define weight gain or loss

wideCW <- dcast(ChickWeight, Diet + Chick ~ Time, value.var = "weight")
names(wideCW)[-(1:2)] <- paste("time", names(wideCW)[-(1:2)], sep = "")
library(dplyr)
wideCW <- mutate(wideCW,
                 gain = time21 - time0
                 )
```

Unequal variance T test comparing diets 1 and 4

```{r}
wideCW14 <- subset(wideCW, Diet %in% c(1, 4))
t.test(gain ~ Diet, paired = FALSE, 
       var.equal = TRUE, data = wideCW14)
```

Exact binomial test

Recall this problem, Suppose a friend has 8 children, 7 of which are girls and none are twins

Perform the relevant hypothesis test. 
H0 : p = 0.5 
Ha : p > 0.5
What is the relevant rejection region so that the probability of rejecting is (less than) 5%?

```{r}
pbinom(-1, size = 8, p = .5, lower.tail = FALSE)
pbinom( 0, size = 8, p = .5, lower.tail = FALSE)
pbinom( 1, size = 8, p = .5, lower.tail = FALSE)
pbinom( 2, size = 8, p = .5, lower.tail = FALSE)
pbinom( 3, size = 8, p = .5, lower.tail = FALSE)
pbinom( 4, size = 8, p = .5, lower.tail = FALSE)
pbinom( 5, size = 8, p = .5, lower.tail = FALSE)
pbinom( 6, size = 8, p = .5, lower.tail = FALSE)
pbinom( 7, size = 8, p = .5, lower.tail = FALSE)
```

# P-Values

Most common measure of statistical significance
Suppose nothing is going on - how unusual is it to see the estimate we got?

Approach:

Define the hypothetical distribution of a data summary (statistic) when "nothing is going on" (null hypothesis)
Calculate the summary/statistic with the data we have (test statistic)
Compare what we calculated to our hypothetical distribution and see if the value is "extreme" (p-value)

P-values

The P-value is the probability under the null hypothesis of obtaining evidence as extreme or more extreme than that obtained
If the P-value is small, then either H0 is true and we have observed a rare event or H0 is false
Suppos that you get a T statistic of 2.5 for 15 df testing H0: mu = mu0 versus Ha : mu > mu_0.
What's the probability of getting a T statistic as large as 2.5?

```{r}
pt(2.5, 15, lower.tail = FALSE) #1.2%
# Therefore, the probability of seeing evidence as extreme or more extreme than that actually obtained under

# Either the null hypothesis is true and we saw a large statistics or the null hypothesis is false
```

Revisiting an earlier example

Suppose a friend has 8 children, 7 of which are girls and none are twins
If each gender has an independent 50% probability for each birth, what's the probability of getting 7 or more girls out of 8 births?

```{r}
# H0 = p = 0.5
# Ha = p > 0.5
choose(8,7)*0.5^8 + choose(8,8)*0.5^8 

pbinom(6, size = 8, prob = 0.5, lower.tail = FALSE)
```

Poisson example

Suppose that a hospital has an infection rate of 10 infections per 100 person/days at risk (rate of 0.1) during the last monitoring period.
Assume that an infection rate of 0.05 is an important benchmark.
Given the model, could the observed rate being larger than 0.05 be attributed to chance?
Under H0: lambda = 0.05 so that lambda_0 100 = 5
Consider H_a: lambda > 0.05.

```{r}
ppois(9,5, lower.tail = FALSE)
```