# Multiple Comparisons

## Key ideas

Hypothesis testing/significance analysis is commonly overused
Correcting for multiple testing avoids false positives or discoveries
Two key components
Error measure
Correction

Case study I: no true positives

```{r}
set.seed(1010093)
pValues <- rep(NA,1000)
for(i in 1:1000){
  y <- rnorm(20)
  x <- rnorm(20)
  pValues[i] <- summary(lm(y ~ x))$coeff[2,4]
}

# Controls false positive rate
sum(pValues < 0.05)

# Case study I: no true positives

# Controls FWER 
sum(p.adjust(pValues,method="bonferroni") < 0.05)
# Controls FDR 
sum(p.adjust(pValues,method="BH") < 0.05)
```

Case study II: 50% true positives

```{r}
set.seed(1010093)
pValues <- rep(NA,1000)
for(i in 1:1000){
  x <- rnorm(20)
  # First 500 beta=0, last 500 beta=2
  if(i <= 500){y <- rnorm(20)}else{ y <- rnorm(20,mean=2*x)}
  pValues[i] <- summary(lm(y ~ x))$coeff[2,4]
}
trueStatus <- rep(c("zero","not zero"),each=500)
table(pValues < 0.05, trueStatus)

# Case study II: 50% true positives

# Controls FWER 
table(p.adjust(pValues,method="bonferroni") < 0.05,trueStatus)
# Controls FDR 
table(p.adjust(pValues,method="BH") < 0.05,trueStatus)

# Case study II: 50% true positives

# P-values versus adjusted P-values

par(mfrow=c(1,2))
plot(pValues,p.adjust(pValues,method="bonferroni"),pch=19)
plot(pValues,p.adjust(pValues,method="BH"),pch=19)
```

Notes and resources

Notes:

Multiple testing is an entire subfield
A basic Bonferroni/BH correction is usually enough
If there is strong dependence between tests there may be problems
Consider method="BY"