# Quiz 2 - Statistical Inference
1. What is the variance of the distribution of the average an IID draw of n observations from a population with mean ?? and variance ??2.

??2 / n

2. Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?

```{r 2}

pnorm(70, mean = 80, sd = 10)

```

3. Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?

```{r 3}

qnorm(0.95, mean = 1100, sd = 75)

```

4. Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?

```{r 4}

qnorm(0.95, mean = 1100, sd = 75/sqrt(100))

```

5. You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?

```{r 5}

pbinom(3, size = 5, prob = 0.5, lower.tail = FALSE)

```

6. The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?

```{r 6}

mu <- 15
si <- 10
n <- 100
s <- si/sqrt(n)

probs <- (pnorm(16, mean = mu, sd = s)-pnorm(14, mean = mu, sd = s))
probs


```

7. Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?

```{r 7}

mu <- 0.5
si <- 1/12
n <- 1000
s <- si/sqrt(n)

qnorm(0.5, mean = mu, sd = s)


```

8. The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?

```{r 8}

ppois(10, lambda = 5*3)


```