---
title: "Developing Data Products Course Project: Shiny Application and Product Pitch"
author: Marcelo M Nobrega
date: "August 9th, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This presentation was built as a part of the deliverable for the Final Course Project, as a part of the Coursera Data Science Specilization.

The App developed is available at: <https://marcelomedre87.shinyapps.io/Shiny_Application/>.

Source code for ui.R and server.R files are avilable on the GitHub repo: <https://github.com/marcelomedre/Data-Science-Specialization/tree/master/Data%20products>.

## App Functionality

The App compares the expenditure of a car using gasoline and ethanol given the price of each fuel and the distance of the trip.

The result is a table ordered by the more economic cars. 

The App was build with the mtcars data set. 

## Data set

Data set used was the `mtcars` from the datasets package, which is summarized below:

```{r cars, echo = TRUE}
summary(cars)
```

## Ethanol x Gasoline x Displacement
        - Blue dots = Gasoline Expenditures
        - Red dots = Ethanol Expenditures costing 70% of the gasoline.
        - Salmon dots = Ethanol Expenditures costing 60% of the gasoline.

```{r cars2, echo = FALSE, fig.width=6, fig.height=4}
dist <- seq(0,100,10)
costg <- dist/10*3.65
costet7 <- dist/7*2.55
costet6 <- dist/7*2.19
library(ggplot2)
ggplot()+
        geom_line(aes(x = dist, y = costg), colour = "blue")+
        geom_line(aes(x = dist, y = costet7), colour = "red")+
        geom_line(aes(x = dist, y = costet6), colour = "salmon")+
        xlab("Distance")+
        ylab("Cost")+
        ggtitle("Comparison of a 0 - 100 km trip using Gasoline and Ethanol")
```

