# Basic idea

  - The R function creates an HTML page  
  - The HTML page calls Google Charts 
  - The result is an interactive HTML graphic  

```{r gv, results = "asis", cache = TRUE}
suppressPackageStartupMessages(library(googleVis))
M <- gvisMotionChart(Fruits, "Fruit", "Year",
                     options=list(width=600, height=400))
plot(M)
print(M,"chart")
```

## Charts in googleVis

<center> "gvis + CharType" </center>

  *Motion charts: gvisMotionChart  
  *Interactive maps: gvisGeoChart  
  *Interactive tables: gvisTable  
  *Line charts: gvisLineChart  
  *Bar charts: gvisColumnChart
  *Tree maps: gvisTreeMap  
http://cran.r-project.org/web/packages/googleVis/googleVis.pdf


## Plots on maps
```{r}
G <- gvisGeoChart(Exports, locationvar="Country",
                  colorvar="Profit",options=list(width=600, height=400))
print(G,"chart")
```

## Specifying a region
```{r}
G2 <- gvisGeoChart(Exports, locationvar="Country",
                  colorvar="Profit",options=list(width=600, height=400,region="150"))
print(G2,"chart")
```

## Setting more options

```{r}
df <- data.frame(label=c("US", "GB", "BR"), val1=c(1,3,4), val2=c(23,12,32))
Line <- gvisLineChart(df, xvar="label", yvar=c("val1","val2"),
        options=list(title="Hello World", legend="bottom",
                titleTextStyle="{color:'red', fontSize:18}",                         
                vAxis="{gridlines:{color:'red', count:3}}",
                hAxis="{title:'My Label', titleTextStyle:{color:'blue'}}",
                series="[{color:'green', targetAxisIndex: 0}, 
                         {color: 'blue',targetAxisIndex:1}]",
                vAxes="[{title:'Value 1 (%)', format:'##,######%'}, 
                                  {title:'Value 2 (\U00A3)'}]",                          
                curveType="function", width=500, height=300                         
                ))
```

https://github.com/mages/Introduction_to_googleVis/blob/gh-pages/index.Rmd

## Setting more options

```{r}
print(Line,"chart")
```

##Combining multiple plots together

```{r}
G <- gvisGeoChart(Exports, "Country", "Profit",options=list(width=200, height=100))
T1 <- gvisTable(Exports,options=list(width=200, height=270))
M <- gvisMotionChart(Fruits, "Fruit", "Year", options=list(width=400, height=370))
GT <- gvisMerge(G,T1, horizontal=FALSE)
GTM <- gvisMerge(GT, M, horizontal=TRUE,tableOptions="bgcolor=\"#CCCCCC\" cellspacing=10")
```

## Combining multiple plots together

```{r}
print(GTM,"chart")
```

## Seeing the HTML code

```{r}
M <- gvisMotionChart(Fruits, "Fruit", "Year", options=list(width=600, height=400))
print(M)
print(M, 'chart', file='myfilename.html')
```

## Things you can do with Google Vis

  - The visualizations can be embedded in websites with HTML code  
  - Dynamic visualizations can be built with Shiny, Rook, and R.rsp  
  - Embed them in R markdown based documents  
  - Set results="asis" in the chunk options  
  - Can be used with knitr and slidify  