#Plotly

A basic scatterplot is easy to make, with the added benefit of tooltips that apear when your mouse hovers over each point. Specify a scatterplot by indicating 'mode = "markers"'.

```{r, eval = FALSE}
library(plotly)
plot_ly(mtcars, x = mtcars$wt, y = mtcars$mpg, mode = "markers")
```

## Basic Scatterplot

```{r, echo = FALSE, message = FALSE}
plot_ly(mtcars, x = mtcars$wt, y = mtcars$mpg, mode = "markers", color = as.factor(mtcars$cyl))
```

## Continuous Color

You can also show continuos variables with color in scatterplots

```{r, echo = FALSE, message = FALSE}
plot_ly(mtcars, x = mtcars$wt, y = mtcars$mpg, mode = "markers", color = mtcars$disp)
```

## Scatterplot Sizing

```{r, echo = FALSE, message = FALSE}
plot_ly(mtcars, x = mtcars$wt, y = mtcars$mpg, mode = "markers", color = as.factor(mtcars$cyl), size = mtcars$hp)
```

## 3D Scatterplot

```{r, eval = FALSE}
set.seed(123)
temp <- rnorm(100, mean = 30, sd = 5)
pressure <- rnorm(100)
dtime <- 1:100
plot_ly(x = temp, y = pressure, z = dtime, type = "scatter3d", mode = "markers", color = temp)
```

## Line Graph

Line graphs are the default graph for 'plot_ly()'. They're of course useful for showing change over time:  

```{r, eval = FALSE}
data("airmiles")
plot_ly(x = time(airmiles), y = airmiles)
```

## Multi Line Graph

```{r, eval = FALSE}
library(plotly)
library(tidyr)
library(dplyr)
data("EuStockMarkets")

stocks <- as.data.frame(EuStockMarkets) %>%
  gather(index, price) %>%
  mutate(time = rep(time(EuStockMarkets), 4))

plot_ly(stocks, x = stocks$time, y = stocks$price, color = stocks$index)
```