# Introduction to Regression Models and Least Squares

## Galton's Data

*Let's look at the data first, used by Francis Galton in 1885.
*Galton was a statistician who invented the term and concepts of regression and correlation, founded the journal Biometrika, and was the cousin of Charles Darwin.
*You may need to run install.packages("UsingR") if the UsingR library is not installed.

*Let's look at the marginal (parents disregarding children and children disregarding parents) distributions first.
  * Parent distribution is all heterosexual couples.
  * Correction for gender via multiplying female heights by 1.08.
  * Overplotting is an issue from discretization.

```{r}
library(UsingR); data(galton); library(reshape); long <- melt(galton)
g <- ggplot(long, aes(x = value, fill = variable)) 
g <- g + geom_histogram(colour = "black", binwidth=1) 
g <- g + facet_grid(. ~ variable)
g
# x-axis = inches, y = counts of children that falls in that inch interval
```

## Finding the middle via least squares

* Consider only the children's heights.
  * How could one describe the "middle"?
  * One definition, let $Y_i$ be the height of child $i$ for $i = 1, \ldots, n = 928$, then define the middle as the value of $\mu$ that minimizes $$\sum_{i=1}^n (Y_i - \mu)^2$$
* This is physical center of mass of the histrogram.
* You might have guessed that the answer $\mu = \bar Y$.

## Experiment
Use R studio's manipulate to see what value of $\mu$ minimizes the sum of the squared deviations.

```{r}
library(manipulate)
myHist <- function(mu){
    mse <- mean((galton$child - mu)^2)
    g <- ggplot(galton, aes(x = child)) + geom_histogram(fill = "salmon", colour = "black", binwidth=1)
    g <- g + geom_vline(xintercept = mu, size = 3)
    g <- g + ggtitle(paste("mu = ", mu, ", MSE = ", round(mse, 2), sep = ""))
    g
}
manipulate(myHist(mu), mu = slider(62, 74, step = 0.5))

# The least squares est. is the empirical mean
g <- ggplot(galton, aes(x = child)) + geom_histogram(fill = "salmon", colour = "black", binwidth=1)
g <- g + geom_vline(xintercept = mean(galton$child), size = 3)
g
```

## Introductory Data Example

Comparing childrens' heights and their parents' heights
```{r}
ggplot(galton, aes(x = parent, y = child)) + geom_point()

```

There is numerous failings in this plot. One of the primary failings is that these points are over-plotted. There's lots of different paired, paired parent child x, y values at each one of these specific points. 

Size of point represents number of points at that (X, Y) combination
```{r}
library(dplyr)
freqData <- as.data.frame(table(galton$child, galton$parent))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
g <- g  + scale_size(range = c(2, 20), guide = "none" )
g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
g <- g + geom_point(aes(colour=freq, size = freq))
g <- g + scale_colour_gradient(low = "lightblue", high="white")                    
g
```

## Regression through the origin
Suppose that we want to explain children's heights with parents heights.

Suppose that $X_i$ are the parents' heights.
Consider picking the slope $\beta$ that minimizes $$\sum_{i=1}^n (Y_i - X_i \beta)^2$$
This is exactly using the origin as a pivot point picking the line that minimizes the sum of the squared vertical distances of the points to the line
Use R studio's manipulate function to experiment
Subtract the means so that the origin is the mean of the parent and children's heights

```{r}
y <- galton$child - mean(galton$child)
x <- galton$parent - mean(galton$parent)
freqData <- as.data.frame(table(x, y))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
myPlot <- function(beta){
    g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
    g <- g  + scale_size(range = c(2, 20), guide = "none" )
    g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
    g <- g + geom_point(aes(colour=freq, size = freq))
    g <- g + scale_colour_gradient(low = "lightblue", high="white")                     
    g <- g + geom_abline(intercept = 0, slope = beta, size = 3)
    mse <- mean( (y - beta * x) ^2 )
    g <- g + ggtitle(paste("beta = ", beta, "mse = ", round(mse, 3)))
    g
}
manipulate(myPlot(beta), beta = slider(0.6, 1.2, step = 0.02))
```

there is a point in regression to the origin is useful for explaining things, because we only have one parameter, the slope and we don't have two parameters, the slope and the intercept. But it's generally bad practice to force regression lines through the point zero, zero. So, an easy way around this is to subtract the mean from the parent's heights and the mean from the child's heights, so that the zero, zero point is right in the middle of the data and that will make this solution a little bit more palatable.

## The Solution

```{r}
lm(I(child - mean(child))~ I(parent - mean(parent)) - 1, data = galton)

```


```{r}
freqData <- as.data.frame(table(galton$child, galton$parent))
names(freqData) <- c("child", "parent", "freq")
freqData$child <- as.numeric(as.character(freqData$child))
freqData$parent <- as.numeric(as.character(freqData$parent))
g <- ggplot(filter(freqData, freq > 0), aes(x = parent, y = child))
g <- g  + scale_size(range = c(2, 20), guide = "none" )
g <- g + geom_point(colour="grey50", aes(size = freq+20, show_guide = FALSE))
g <- g + geom_point(aes(colour=freq, size = freq))
g <- g + scale_colour_gradient(low = "lightblue", high="white")                    
lm1 <- lm(galton$child ~ galton$parent)
g <- g + geom_abline(intercept = coef(lm1)[1], slope = coef(lm1)[2], size = 3, colour = grey(.5))
g

```