# Regression Models - Quiz 3

1. Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

```{r, echo = FALSE}
data("mtcars")
model <- lm(mpg ~ factor(cyl)  + wt, data = mtcars)
summary(model)$coef
```

2. Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?.

```{r, echo = FALSE}
data("mtcars")
modelun <- lm(mpg ~ factor(cyl), data = mtcars)
modeladj <- lm(mpg ~ factor(cyl)  + wt, data = mtcars)
summary(modelun)$coef
summary(modeladj)$coef
```

3.Consider the mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.

```{r, echo = FALSE}
data("mtcars")
model1 <- lm(mpg ~ factor(cyl)  + wt, data = mtcars)
model1$coef
model2 <- lm(mpg ~ factor(cyl) * wt, data = mtcars)
model2$coef
anova(model1, model2)
```
4. Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars) How is the wt coefficient interpretted?

```{r}
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
```

5. Consider the following data set
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
Give the hat diagonal for the most influential point

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit <- lm(y ~ x)
max(hatvalues(fit))
```
6. Consider the following data set
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
Give the slope dfbeta for the point with the highest hat value.

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit <- lm(y ~ x)
influence.measures(fit)$infmat[5, 'dfb.x']
```
7. Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.

