---
output:
  pdf_document: default
  html_document: default
---
# Regression Models Course Project
## Instructions
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

 - "Is an automatic or manual transmission better for MPG"  
 - "Quantify the MPG difference between automatic and manual transmissions"
 
### Load and check data
```{r}
rm(list = ls())
cardata <- mtcars
head(cardata)
summary(cardata) # no NA's
```

### "Is an automatic or manual transmission better for MPG?"  
### Plotting 
```{r}
library(ggplot2)
ggplot(data = cardata, aes(y = mpg, x = factor(am), fill = factor(am)))+
  geom_boxplot()+
  scale_fill_discrete(name="Tranmission\nType")+
  xlab("Transmission type 0 = auto and 1 = manual")+
  ylab("Miles / Gallon")+
  ggtitle("Figure 1 - is automatic or manual transmission better for MPG?")

```

Figure 1 indicates that manual transmission is better for MPG than automatic transmission.
We could check if this diference is significant.

### Checking the significance of the observed MPG diferences

Let's consider:  
  - Null hypothesis: the MPG of the automatic and manual transmission
  has no difference.   
 - Alternative hypothesis: there difference of the MPG for the automatic and manual transmissions
 is significant.  

The hypothesis can be test with a t-test

```{r}
t.res <- t.test(mpg ~ factor(am), data = cardata)
t.res
```

p-value is smaller than 0.05, so the null hypothesis is rejected and it can be said
that there is a significant difference between for the MPG of the automatic and manual cars. 
The mean MPG for cars with automatic transmission is 17.15, while for cars with manual
transmission is 24.39 miles per gallon.

### "Quantify the MPG difference between automatic and manual transmissions"
#### Regression Models

```{r}
linReg <- lm(mpg ~ factor(am), data = cardata)
summary(linReg)
```
Using a linear model to quantify the MPG difference between automatic and manual transmissions
we obtain that automatic cars have 17.15 average MPG, whereas manual ones have 7.24 MPG higher.
Additionally, despite the low p-values, the adj-R-squared is 0.33, which corresponds to a model
that is accounting for only ~33 % of the MPG variance, a very poor model. 

### Improving the model

We can try to improve the model by using all the variables.
```{r}
fullLinReg <- lm(mpg ~. , data = cardata)
summary(fullLinReg)
# plot(residuals(fullLinReg))
```
Using all the variables in the model we got an adj-R-Squared of 0.81, however,
p-values are all greater than 0.05 which indicates no significance of the variables to
the model. We must use only significant variables in the model, which can be done
with the step function. 

### Selecting significant variables
```{r}
stepmodel <- step(lm(mpg ~. , data = cardata), trace = 0, steps = 1000)
summary(stepmodel)
```

With the selected significant variables (wt, qsec and am) the model accounted to the 
explanation of ~ 83% of the MPF variance. 
Am coef indicates that manual transmission cars have 2.9358 more mpg than automatic transmission
cars on average when all other variables are constant.

### Analysis and Diagonostics of Residuals

```{r}
# Residual plot
par(mfrow = c(2,2))
plot(stepmodel)
```

 - In the Residuals vs Fitted values plot no pattern can be recognized. The points are
 scattered around zero.   
 - Normal Q-Q plot most part of the data points fall on the normal distribution line.   
 - Scale-Location indicates constant variance with the scattered points pattern.  
 - Residuals vs Leverage shows that all values are in the 0.5 range indicating that
 no outliers are present in the dataset.
 
## Conclusion

MPG analysis showed that manual transmission cars are better in terms of fuel consumption
when compared to automatic cars with and average MPG of 24.39 for manual cars, 7.24 MPG
higher than the average of automatic cars. 

