# Bagging

Bootstrap Aggregating

The basic idea is that when you fit complicated models, sometimes if you average those models together, you get a smoother model fit, that gives you a better balance between potential bias in your fit and variance in your fit. 

## Basic idea
  - Resample cases and recalculate predictions  
  - Average or majority vote    
  
## Notes 
  - Similar bias  
  - Reduced variance  
  - More Useful for Non-Linear Functions  
  
```{r}
rm(list = ls())
library(ElemStatLearn)
data(ozone)
ozone <- ozone[order(ozone$ozone),]
head(ozone)
```

## Predict Temperature as a function of Ozone

Bagged Loess
Loess = smooth curve fitted to the data

```{r}
ll <- matrix(NA, nrow = 10, ncol = 155)
for(i in 1:10){
  ss <- sample(1:dim(ozone)[1], replace = TRUE)
  ozone0 <- ozone[ss,]
  ozone0 <- ozone0[order(ozone0$ozone),]
  loess0 <- loess(temperature ~ ozone, data = ozone0, span = 0.2)
  ll[i,] <- predict(loess0, newdata = data.frame(ozone = 1:155))
}
plot(ozone$ozone, ozone$temperature, pch = 19, cex = 0.5)
for(i in 1:10) {lines(1:155, ll[i,], col = "grey", lwd = 2)}
lines(1:155, apply(ll,2,mean), col = "red", lwd = 2)
```

## Bagging in caret

Using the train function
  - bagEarth
  - treebag
  - bagFDA
  
Own Bagging Function
```{r}
library(caret)
predictors <- data.frame(ozone = ozone$ozone)
temperature <- ozone$temperature
treebag <- bag(predictors, temperature, B = 100,
               bagControl = bagControl(fit = ctreeBag$fit,
                                       predict = ctreeBag$pred,
                                       aggregate = ctreeBag$aggregate))
plot(ozone$ozone, temperature, col = "lightgrey", pch = 19)
points(ozone$ozone, predict(treebag$fits[[1]]$fit, predictors), pch = 19, col = "red")
points(ozone$ozone, predict(treebag, predictors), pch = 19, col = "blue")
```

## Parts of bagging
```{r}
ctreeBag$fit
ctreeBag$pred
ctreeBag$aggregate
```

 - Bagging is most useful for nonlinear models  
 - Often used with trees - an extension is random forests  
 - Several models use bagging in caret's train function  



