# Forecasting

## What is different?

  - Data are dependent over time  
  - Specific pattern types  
    - Trends - long term increase or decrease  
    - Seasonal patterns - patterns related to time of week, month, year, etc.  
    - Cycles - patterns that rise and fall periodically  
  - Subsampling into training/test is more complicated  
  - Similar issues arise in spatial data  
    - Dependency between nearby observations    
    - Location specific effects   
  - Typically goal is to predict one or more observations into the future.  
  - All standard predictions can be used (with caution!)  

## Example Google Data
```{r}
library(quantmod)
from.dat <- as.Date("01/01/08", format = "%m/%d/%y")
to.dat <- as.Date("12/31/13", format = "%m/%d/%y")
getSymbols("GOOG", src = "google", from = from.dat, to = to.dat)
head(GOOG)
```

##Summarize monthly and store as time series

```{r}
mGoog <- to.monthly(GOOG)
googOpen <- Op(mGoog)
ts1 <- ts(googOpen,frequency=12)
plot(ts1,xlab="Years+1", ylab="GOOG")
```

## Example time series decomposition

Trend - Consistently increasing pattern over time  
Seasonal - When there is a pattern over a fixed period of time that recurs.  
Cyclic - When data rises and falls over non fixed periods  
https://www.otexts.org/fpp/6/1  

## Decompose a time series into parts
```{r}
plot(decompose(ts1),xlab="Years+1")
```

## Training and test sets
```{r}
ts1Train <- window(ts1,start=1,end=5)
ts1Test <- window(ts1,start=5,end=(7-0.01))
ts1Train
```

## Simple moving average

$$ Y_{t}=\frac{1}{2*k+1}\sum_{j=-k}^k {y_{t+j}}$$
```{r}
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")
```

## Exponential smoothing
```{r}
ets1 <- ets(ts1Train,model="MMM")
fcast <- forecast(ets1)
plot(fcast); lines(ts1Test,col="red")
```

## Get the accuracy
```{r}
accuracy(fcast,ts1Test)
```

## Notes and further resources

Forecasting and timeseries prediction is an entire field  
Rob Hyndman's Forecasting: principles and practice is a good place to start  
Cautions  
  - Be wary of spurious correlations  
  - Be careful how far you predict (extrapolation)  
  - Be wary of dependencies over time  
See quantmod or quandl packages for finance-related problems.