# Predicting with Regression

## Key ideas
  - Fit a single regression model  
  - Plug in new covariates and multiple by coeff  
  - Useful when the linear model is nearly correct

Pros  
  - Easy to implement  
  - Easy to interpret

Cons  
  - Often poor performance in nonlinear settings

Example:
Data on eruptions of geysers
```{r}
library(caret)
data("faithful")
set.seed(333)
inTrain <- createDataPartition(y = faithful$waiting, p = 0.5, list = FALSE)
trainFaith <- faithful[inTrain,]
testFaith <- faithful[-inTrain,]
head(trainFaith)
```

```{r}
plot(trainFaith$waiting, trainFaith$eruptions, pch = 19, col = "blue", xlab = "Waiting", ylab = "Duration")
```

Fit a linear Model
```{r}
lm1 <- lm(eruptions ~ waiting, data = trainFaith)
summary(lm1)
```

```{r}
plot(trainFaith$waiting, trainFaith$eruptions, pch = 19, col = "blue", xlab = "Waiting", ylab = "Duration")
lines(trainFaith$waiting, lm1$fitted.values, lwd = 3)
```

## Predict a new value
```{r}
coef(lm1)[1] + coef(lm1)[2]*80
newdata <- data.frame(waiting = 80)
predict(lm1, newdata)
```

## Plot predictions - training and test

```{r}
par(mfrow = c(1,2))
plot(trainFaith$waiting, trainFaith$eruptions, pch = 19, col = "blue", xlab = "Waiting", ylab = "Duration")
lines(trainFaith$waiting, lm1$fitted.values, lwd = 3)
plot(testFaith$waiting, testFaith$eruptions, pch = 19, col = "red", xlab = "Waiting", ylab = "Duration")
lines(testFaith$waiting, predict(lm1, newdata = testFaith), lwd = 3)
```

## Get training set/test set errors

```{r}
# Calculate RMSE train
sqrt(sum((lm1$fitted.values - trainFaith$eruptions)^2))

# Calculate RMSE on test
sqrt(sum((predict(lm1, newdata = testFaith) - testFaith$eruptions)^2))
```

## Prediction intervals

```{r}
pred1 <- predict(lm1, newdata = testFaith, interval = "prediction")
ord <- order(testFaith$waiting)
plot(testFaith$waiting, testFaith$eruptions, pch = 19, col = "blue", xlab = "Waiting", ylab = "Duration")
matlines(testFaith$waiting[ord], pred1[ord,], type = "l", col = c(1,2,2), lty = c(1,1,1), lwd = 3)
```

## Same process with carer

```{r}
modFit <- train(eruptions ~ waiting, data = trainFaith, method = "lm")
summary(modFit)
```