# Unsupervised prediction

## Key ideas

  - Sometimes you don't know the labels for prediction  
  - To build a predictor  
    - Create clusters  
    - Name clusters  
    - Build predictor for clusters  
  - In a new data set  
  - Predict clusters  
  
## Iris example ignoring species labels

```{r}
data(iris); library(ggplot2)
inTrain <- createDataPartition(y=iris$Species,
                              p=0.7, list=FALSE)
training <- iris[inTrain,]
testing <- iris[-inTrain,]
dim(training); dim(testing)
```

## Cluster with k-means

```{r}
kMeans1 <- kmeans(subset(training,select=-c(Species)),centers=3)
training$clusters <- as.factor(kMeans1$cluster)
qplot(Petal.Width,Petal.Length,colour=clusters,data=training)
```

## Compare to real labels
```{r}
table(kMeans1$cluster,training$Species)
```

## Build predictor
```{r}
modFit <- train(clusters ~.,data=subset(training,select=-c(Species)),method="rpart")
table(predict(modFit,training),training$Species)
```

## Apply on test
```{r}
testClusterPred <- predict(modFit,testing) 
table(testClusterPred ,testing$Species)
```

## Notes and further reading

The cl_predict function in the clue package provides similar functionality  
Beware over-interpretation of clusters!  
This is one basic approach to recommendation engines  
Elements of statistical learning  
Introduction to statistical learning  